Distrivia: A Distributed Trivia Game

### 2011-04-06 papers/phase-one/paper.tex
Setup and add phase-one document latex paper...

\section{Overview}
Our idea for the P3 project is a trivia game built for a distributed system, which we're calling Distrivia.
The premise of the program is to deliver multiple-choice trivia questions to the players in the form of rounds.
Ten questions will be delivered for each round.
Players are connected up in groups of 20 players, with each player's score dependent on the speed in which they answer each question.
Once each player has completed the round, the leader board for the round will display the scores for each player for that round.
After which, the players will join new rounds to compete against other players.
Distrivia will also keep a tally of their global score (possibly an average), which will determine how well the player competes on a whole.
This global score could give the player a unique symbol after their name on leader boards or perhaps a distinct color.

The trivia questions will come from many different aspects of academia and at varying level of difficulty
This will assure that there is an equal chance for players to get questions correct and compete.

We would also like to implement a way for players to create a private game so they can compete with friends.
This will be accomplished by the initial player naming the game, and the other players joining the game with the name.
Private games allow users to compete with people they know which will allow the game to be more interactive.

\section{Implementation}
We will be creating software, initially, for desktops, laptops, and android phones.
The software for desktops/laptops will be implemented in a web browser.
If there is time to implement it, we want to create an application for the iPhone as well.
The applications for all platforms will look identical and allow the players to have the same functionality, independent of which device they are using to play Distrivia.

%\subsection{Clients}
The client will be pretty straightforward.
If you have never played Distrivia before, the game will prompt for the player to create a player name.
After which, it will allow the user to request to play a game.
Once the server has grouped 20 players together, the round will start and the screen will show the question followed by 4 selectable answers.
Once the user has selected their answer, they can hit the Submit button at the bottom to submit their answer for that question.
At that point, the player's answer is sent to the server, their new score along with their next question is sent back to the player.
At the bottom of the screen, the player will be able to select from thumbs up/thumbs down buttons if they particularly like or dislike a question.
This will be particularly useful if we have time to allow users to create their own questions to add the to the database.
Once a questions dislike ratio is too high, the question will no longer be asked.

  \includegraphics[scale=0.3]{design-pictures/login.png}
  \includegraphics[scale=0.3]{design-pictures/joingame.png}
  \includegraphics[scale=0.3]{design-pictures/quiz.png}
  \includegraphics[scale=0.3]{design-pictures/leaderboard.png}

%\subsection{Backend}
The server has four key layers involved.
First, there will be a load balancer that accepts all messages from the players.
It does not need to read the contents of the message, it simply distributes the message on to the most reliable server with the lowest load.
This ensures that if any of the servers were to go down, that the system would still run.
It also maintains highest speeds for server response.
The message is passed to a web server from the load balancer, which will read the contents of the message and respond appropriately.
These messages will be in the form of http messages.
The web server communicates directly with the database to perform whatever actions need to be done.
The third layer is represented by these synchronous databases that hold all the actual data for the software.
This would include round information, player information, and all the questions for the application.
Currently we are looking at using Riak for our database engine.
Riak has the unique ability to perform automatic master less replication and
synchronization between all nodes of a database cluster. With this architecture 
all back end nodes except for one could fail and our system would remain
operational.
\begin{center}
  \includegraphics[scale=0.35]{design-pictures/database2.png}
\end{center}

>>>>

\section{What We've Done}
So far, our work has mainly focused on getting everything setup to work as a
distributed system. Getting the ground work done (for example servers and load
distributers running) has provided us a way to quickly develop the clients and
server as well as provided us with a way to reliably connect the servers to the
clients.

\subsection{Servers}
Currently, we have 4 servers running. One server is being used as a web-monitor
host. We are able to login to the web-monitor to see the status of all our
servers, databases, and load distributers. This will help us manage if the
services are running and allow us to track down any problems, if they arise.
The other three servers are all running our server application and hosting the
databases. Each server contains the server application that responds to clients
as well as the Riak databases. Riak auto-synchronizes the databases between the
3 servers. If we were to add a server, we would just need to updatrver as well
as provided us with a way to reliably connect the servers to the clients.

\subsection {Traffic}
We have two load distributers set up. Both distributers are currently
configured to divide the requests among the 3 active servers. The duplicate
distributers allow for a load distributer to go down and we would still be able
to serve requests to the servers. These load distributers are selected by the
clients using round robin DNS where distrivia.lame.ws will select either load
distributer that is up at random. This provides us will stability for access to
the servers. The load distributers are set up the attempt to connect to a
server twice before deeming it dead and removing it from its list of servers to
direct clients towards.

\subsection {Clients}
Our focus for clients has been on a web-based client and an android
application. Both clients have most of the their screens completed. The
web-based client will be run from desktop platforms and laptop platforms. The
android client will run on any android phone and does not use any unique
services that require specific hardware.

\section{What's Left}
\subsection{Web Client}
Joining/playing private games
In-game leaderboard
Finish connecting to actual serversver API
Implement useful user warnings

\subsection{Android}
Connecting to server API
Registration
Private games
Leaderboards
Allowing users to control screens

\subsection{Server}
Passwords
Private game joining
Global leaderboard
Finish setting up SSL on load distributers
Set up 2 hot spares

\section {Meeting the Requirements}
-3 Devices (Android, Desktop, Laptop)
-Reliable / Flexible communication (Multiple servers, multiple load
distributers, 2 hot spares, proper communication to user)
-Coordination/Cooperation (Synchronous Databases and Rounds consisting of many
players)
-Replication / Availability / Consistency (Riak handles synchronizing databases
among servers. Load distributers provide availability)
- Naming / Service Discovery
- Fault tolerance / Handling of partial failures
- Security (SSL)
- Scalability (add new nodes and add to riak cluster and it's all set)
- Use of an interesting distributed algorithm
